%============================= abstract.tex================================
\chapter*{Abstract}%
%\addcontentsline{toc}{chapter}{\numberline{}Abstract}%
\addcontentsline{toc}{chapter}{Abstract}%
The MealMingle app is a social initiative that addresses food waste and food insecurity by connecting individuals, hotels, restaurants, and NGOs. Users can create profiles, browse listings, and request meals based on their preferences. Hotels and restaurants can contribute surplus meals to the app, which are made available for users and NGOs to request. The app streamlines the process by allowing NGOs to directly book meals for individuals in need, ensuring efficient distribution. MealMingle fosters community collaboration and compassion, promoting a sustainable approach to food consumption.

By sharing meals, donating to hunger relief charities, and volunteering, users actively contribute to reducing food waste and alleviating hunger. The app provides an inclusive platform for individuals to connect and support each other. MealMingle's user-friendly interface facilitates a seamless experience, allowing users to make a meaningful impact and create positive change in their communities. Together, we can foster empathy, unity, and social responsibility, making a difference in the lives of the underserved.

In summary, MealMingle is a platform that connects individuals, hotels, restaurants, and NGOs to address food waste and food insecurity. Users can browse and request meals, while hotels and restaurants can contribute surplus meals. NGOs have the ability to directly book meals for individuals in need. MealMingle promotes community collaboration and offers a user-friendly interface for sharing meals, donating, and volunteering. By actively participating in the app, users contribute to reducing food waste, alleviating hunger, and fostering a stronger sense of social responsibility.
\thispagestyle{plain}
%=======================================================================

